{\it Maxwell's Silver Hammer:}\\ On pulsed microwave acoustic resonant viral inactivation

### 2021-02-09 documents/background.tex
Text and prop improvements

@@

%3 MV/m * 10000 electronCharge * 0.1 angstrom / (boltzmannConstant * 310 K) = 11 kT per cycle
%3 MV/m * 10000 electronCharge * 0.1 angstrom / (boltzmannConstant * 310 K) = 11 kT per cycle
From an energetics perspective, we can obtain the spring constant $k = \omega_{res}^2 m^*$ from the resonant frequency, where $m^*$ is the reduced mass. With relatively sane choices of screened charge and resonant frequency, the energy per monocycle at 5 MV/m varies from 0.01 to 20 times the thermal energy $k_b T$.

Compare to ~5 $k_b T$ per protein for capsid binding \cite{Energies2012} \cite{Weak2002} and $ (\text{17 to 78 kJ/mol})  / (310 \text{ K}\ k_b) = $ 6 to 30 $k_b T$ for lipid bilayer pore formation \cite{Atomistic2014a}.
%17 or 78 kJ/mol / (boltzmannConstant * 310 K) / avogadroConstant = 6.5 to 30.3 kT

@@

\begin{autem}
Extreme care must be taken in interpreting the results of any one study in isolation, especially in this field.

Though the temperature was monitored, In this observation, the thermal artifact cannot necessarily be completely discounted. The energy dissipated was sufficient to cause a significant temperature rise if localized hot-spots appeared \footnote{(43 cal/cc * (3 microliters)) / (3 milligrams * 4.2 J/(g K)) = ~50 $\Delta$C, whereas melting temperatures are typically about 90 C for capsids \cite{Thermal1999} and 50 C for envelopes\cite{Stability1985}}. \\ The electrodes were not insulated, so electrochemical effects are possible but unlikely given the short duration of the pulses. 
\end{autem}

Madiyar \cite{Manipulation2013} observe rapid permeabilization (likely through the slow, conventional electroporation mechanism) at a field of poorly-defined intensity up to about 10 MV/m, in the Vaccinia poxvirus (a 300 nm enveloped dsDNA virus). \cite{AC2017}. Interestingly, the same effect is not observed in T4 bacteriophage.

@@

@@

Man \cite{Picosecond2016b} perform a trustworthy all-atom molecular dynamics simulation of . As with most molecular dynamics simulations, the applied field was enhanced by several orders of magnitude to an unrealistic 2 gigavolts per meter to make the reaction proceed fast enough to be captured without excessive computation time.

@@

A group of literature exists concerning the excitation of these modes with picosecond or femtosecond lasers. A pulse some 300 fs long and with a carrier of around 600 nm. Impulsive stimulated Raman scattering. We confess to not having studied these mechanisms in great detail. The pulse does not directly act on the vibrational mode; the optimal pulse duration is some small fraction of the period of the mode in question.

@@

It is common for. However, non-zero resonant frequency requires the ratio of the the loss in energy to the surroundings to the mass, to be greater than $b/(2m) < \omega_{res} \sqrt{2}$ \cite{Driven}; the solution to the harmonic oscillator equation is piecewise around this value.

@@

We are adamant that, in all cases cited, we believe their studies were well designed and conducted: the authors went to considerable and sufficient lengths to demonstrate its validity. 
\end{toolchain}

@@

@@

@@

@@

A previous report of microwave resonance, Grundler \cite{Sharp1983}, use a calorimeter, a stream of cooling water controlled to within 0.1 C, an interferometric thermometer, and can state definitively that at no point does the temperature inhomogeneity exceed 0.02 C.

@@

Kozempel \cite{Preliminary1997} and \cite{Inactivationa}. They take great pains to map the temperature profile using a fiber-optic thermometer; use thermographic indicators; and find substantial, almost complete inactivation of {\it Pediococcus sp.} with a measured discharge temperature of only 40 C.

This is a fascinating result, and one which should be compared to the existing literature and very carefully replicated to isolate the mechanism. However, unless it is suggested that {\it Pediococcus sp.} also exhibits a confined microwave resonance of the same nature, it appears that periodic temperature measurements within the physiological range are not necessarily a sufficiently stringent control to definitively {\it isolate} resonant effects from other mechanisms.

Conversely, if temperature control is not possible, the temperature can be allowed to rise so long as a positive control is present in which

@@

Burton\cite{Effects1950} describe the use of thermochromic pigments to examine heating profiles. Using low-frequency exposure, they are able to produce localized regions in excess of 50 C.

@@

	To confirm that our observation is not due to the microwave thermal heating effect, we had monitored the sample temperature change during the microwave illumination experiments with a radiated power density of 486W/m2 at a frequency of 6GHz by using an infrared thermal imaging camera with a temperature accuracy of 0.05°C (CHCT, P384-20). The temperature rise after 15 minutes radiation was 7°C, from 27.5°C up to 34.5°C. We thus exclude the possible contribution of microwave thermal heating effect to inactivation H3N2 viruses under our experimental condition.

@@

\subsection{Other artifacts}

@@

Very recently, Burkhartsmeyer \cite{Optical2020} have published data replicating the findings using a single-particle Raman spectroscopy technique. We have not examined these claims in detail, but they are good evidence.

Non-photothermal optical Raman methods can also apparently synthesize artifacts - see unpublished, personal communication with Taylor \cite{mechanisms1981}.

@@

This technique appears to be unique in that it is selective, save for Far UV or cold-plasma sterilization. Its advantages over those is that it operates over large volumes with inexpensive emitters, but, most critically, can be made to penetrate tissue without harm.

>>>>

After the relaxation time, the energy is a distant memory; only the mechanical effects that were induced during that time can remain. Since the energy is too small to

@@

%3 MV/m * 10000 electronCharge * 0.1 angstrom / (boltzmannConstant * 310 K) = 11 kT per cycle
%3 MV/m * 10000 electronCharge * 0.1 angstrom / (boltzmannConstant * 310 K) = 11 kT per cycle
From an energetics perspective, we can obtain the spring constant $k = \omega_{res}^2 m^*$ from the resonant frequency and therefore from the speed of sound and $m^*$, the reduced mass. With relatively sane choices of screened charge and resonant frequency, the energy per monocycle at 5 MV/m varies from 0.01 to 20 times the thermal energy $k_b T$.

Compare to ~5 $k_b T$ per protein for 1 capsid segment binding energy \cite{Energies2012} \cite{Weak2002} and $ (\text{17 to 78 kJ/mol})  / (310 \text{ K}\ k_b) = $ 6 to 30 $k_b T$ to form a single pore in a lipid bilayer \cite{Atomistic2014a}.
%17 or 78 kJ/mol / (boltzmannConstant * 310 K) / avogadroConstant = 6.5 to 30.3 kT

On the other hand, 0.5 angstroms is only a 0.04\% deflection of the capsid, whereas a general value for host cells is about 3\% deflection.

Certainly the virus will be destroyed at some finite electric field. It remains to be seen whether a practical limit

 \cite{Resonance1996}

@@

While this result would seem to perfectly align with our hypothesis, extreme care must be taken in interpreting the results of any one study in isolation, especially in this field.

See the section on pulse artifacts. Electrode insulation is not mentioned, so electrochemical effects are possible but unlikely given the short duration of the pulses. The observed RNA damage is unexpected from a mechanical standpoint but is somewhat like the damage due to copper ions found by \cite{Microwave1987}; therefore electrochemical interaction may have been a confounding factor.

Also, although the temperature was monitored, the thermal artifact cannot necessarily be completely discounted in this experiment. The energy dissipated was sufficient to cause a significant temperature rise if localized hot-spots were to appear \footnote{(43 cal/cc * (3 microliters)) / (3 milligrams * 4.2 J/(g K)) = ~50 $\Delta$C, whereas melting temperatures are typically about 90 C for capsids \cite{Thermal1999} and 50 C for envelopes\cite{Stability1985}}.

Madiyar \cite{Manipulation2013} observe rapid permeabilization (perhaps through the slow, conventional charge electroporation mechanism) at a field of poorly-defined intensity up to about 10 MV/m, in the Vaccinia poxvirus (a 300 nm enveloped dsDNA virus). \cite{AC2017}. Interestingly, the same effect is not observed in T4 bacteriophage.

@@

Schnelle\cite{Trapping1996}

\subsection{Electrorotation}

and Fuhr\cite{Radiofrequency1994} and Grom \cite{Accumulation2006} report on a modification of so-called dielectrophoresis techniques called electrorotation, in a so-called field cage.

Rotation is sub-synchronous at a few Hz, whereas the field rotates at between 0.1 MHz to over 50 MHz; in all cases the frequency is made high enough that

In the virus tests by Schnelle, a field of 0.3 MV/m was used, without apparent damage to the virus, though their

Gimsa \cite{New1999} use a four-electrode cage at 25 kV/m

@@

Man \cite{Picosecond2016b} perform a trustworthy all-atom molecular dynamics simulation of a laser pulse tuned to provoke dissociation of a viral capsid. In this case the large-scale structual modes are not involved. As with many molecular dynamics simulations using applied fields, the field was enhanced by several orders of magnitude to an unrealistic 2 gigavolts per meter to make the reaction proceed fast enough to be captured without excessive computation time.

@@

A group of literature exists concerning the excitation of these structural modes with picosecond or femtosecond lasers. A pulse some 300 fs long with a carrier of around 600 nm. Impulsive stimulated Raman scattering. We confess to not having studied this mechanism in great detail. The pulse does not directly act on the vibrational mode; the optimal pulse duration is some small fraction of the period of the mode in question.

@@

It is common for.  However, in the classical case, a non-zero resonant frequency requires the ratio of the loss in energy to the surroundings to the oscillating mass to obey $b/(2m) < \omega_{res} \sqrt{2}$ \cite{Driven}; the solution to the harmonic oscillator equation is piecewise around this value.

(unless the various small-scale quantum mechanical oscillators are considered, which in our case appear to require fields in excess of 1 TW to access).

@@

We are adamant that, in all cases cited, we believe their studies were well designed and conducted: the authors went to considerable and sufficient lengths to demonstrate its validity. It is easy to say what could have happened, but much harder to say what did happen.
\end{toolchain}

@@

Vijayalaxmi \cite{Biological2016} \cite{Comprehensive2018} \cite{Funding2019} already offer a number of excellent guidelines on best practices in this field; as do Chou \cite{chou1996radio}. We only offer a few issues which we were particularly concerned about when designing our experiments and searching literature.

@@

Very recently, Burkhartsmeyer \cite{Optical2020} have published data replicating the findings on bacteriophage PhiX174 using a single-particle Raman spectroscopy technique. We have not examined these claims in any detail, but believe that they are excellent evidence.

Perhaps these results cannot be trusted implicitly, however, because non-photothermal optical Raman methods can also apparently synthesize artifacts - see unpublished, personal communication with Taylor \cite{mechanisms1981}. As this is an unpublished negative result, we know nothing about the experimental details that led to this error.

@@

mouse was heating

A few reports in the 1940s suggested that bacteria might be non-thermally inactivated. However, Burton\cite{Effects1950} described an experiment using thermochromic pigments to examine heating profiles throughout the sample. With a half-second pulse of an exposure frequency of only 30 MHz, they observe highly localized regions that exceed 50 C even while the maximum temperature measured by thermocouple never exceeds 36-40 C. This is despite the use of a tube only 9 mm across, where it might be thought that thermal conduction would dissipate any .

With the shorter wavelengths of microwave frequencies, it seems likely that such effects will be far more pronounced.

This is highlighted by the experiments of Kozempel \cite{Preliminary1997} and \cite{Inactivationa}, discussing the performance of an RF fluid sterilization unit. They take great pains to map the temperature profile using a fiber-optic thermometer; use thermographic indicators; and yet find substantial inactivation of {\it Pediococcus sp.} with a maximum measured temperature of only 40 C.

This is a fascinating result, and one which should be compared to the existing literature and very carefully replicated to isolate the mechanism. However, unless it is suggested that {\it Pediococcus sp.} also exhibits a confined microwave resonance of the same nature as the virus, it appears that temperature measurements within the physiological range are not necessarily a sufficiently stringent control to definitively {\it isolate} resonant effects from other mechanisms.

The temperature rise after 15 minutes radiation was 27.5°C up to 34.5°C.

@@

A previous report of microwave resonance, Grundler \cite{Sharp1983} use a calorimeter, a stream of cooling water controlled to within 0.1 C, an interferometric thermometer, and can state definitively that at no point does the temperature inhomogeneity exceed 0.02 C.

@@

Conversely, if temperature control is not practical, the temperature can be allowed to rise - so long as a positive control is present which can be

@@

@@

\begin{quote}
	To confirm that our observation is not due to the microwave thermal heating effect, we had monitored the sample temperature change during the microwave illumination experiments with a radiated power density of 486W/m2 at a frequency of 6GHz by using an infrared thermal imaging camera with a temperature accuracy of 0.05°C (CHCT, P384-20).  We thus exclude the possible contribution of microwave thermal heating effect to inactivation H3N2 viruses under our experimental condition.
\end{quote}

@@

\subsection{Even weirder artifacts}

@@

@@

### 2021-02-09 documents/experiment_1.tex
Text and prop improvements

\section{Experiment 1: Attempts at microwave absorption spectroscopy and

\section{Microwave lysis feedback}

@@

At the time, this seemed important to establish, as it would allow the inactivation threshold to be determined without a time-intensive and potentially dangerous infectivity assay step (making it more plausible that we would get biosafety approval on NCoV).

One could imagine it to be otherwise. For instance, the lysis that occurs after phage infection of bacteria does little harm to the cell; only the permeability of the membrane is damaged\cite{GROWTH}. Similarly, the results of the previous studies could perhaps have been obtained without necessarily breaking the virus' envelope.

>>>>

\section{Experiment 1: Attempts at microwave absorption spectroscopy and inactivation in the microsecond pulsed 100 V/m regime}

\subsection{Microwave lysis feedback}

@@

At the time, this seemed important to establish, as it would allow the inactivation threshold to be determined without a time-intensive and potentially dangerous infectivity assay step (making it more plausible that we would get biosafety approval to obtain the microwave spectrum of NCoV).

One could imagine it to be otherwise. For instance, the lysis that occurs after phage infection of bacteria does little harm to the cell; only the permeability of the membrane is damaged\cite{GROWTH}. Similarly, the results of the previous studies could perhaps have been obtained without necessarily destroying the virus' envelope.

### 2021-02-09 documents/pulse_generator.tex
Text and prop improvements

Impedance matching to appears to be a solved problem. \cite{Microchamber2011} provide a good review, and offer a design using a high-impedance exposure cell in parallel with a 50 ohm resistor, which offers excellent performance to at least 1 GHz. Attenuator for diagnostics are also built in.

@@

 Attenuators with such high peak powers are difficult to obtain.

\cite{Characterization2012}

>>>>

Spark gaps

\cite{kHz1995} uses an interesting construction of a pressurized spark gap.

The ZTX417 parts are expensive (\$15 apiece, as of this writing).

[] mention triggering parameters

There seem to be two main designs:

drivers Pockels cells for Q switching.

Schaal patent  which can produce 70 ps edges using less expensive ZTX415 (although the all-important RF "second schematic", the ).

marx or marx-free

@@

Attenuators that are rated for such high peak powers are hard to find.

Oak \cite{fast1991} Figure 1 offers a circuit to evaluate avalanche behavior of transistors. Using this, the

Impedance matching to appears to be a solved problem. \cite{Microchamber2011} provide a good review, and offer a design using a high-impedance exposure cell in parallel with a 50 ohm resistor, which offers excellent performance to at least 1 GHz. Attenuator for diagnostics are also built in. This cell should be used rather than our design. \cite{Characterization2012}
